Horizontal and vertical diversity drive total biomass and its temporal stability in fish communities

### 2020-06-03 manuscript/bef_stability/manuscript.Rmd
Review until data protocol

Biodiversity can be quantified in numerous ways: either as the number of species
(i.e.  species richness) and in number of functional groups, which are usually
defined based on the trophic ecology of species, namely the way they acquire
food [@allan_stream_2007]. Those metrics can be organized into (i) "horizontal
diversity", which captures the overall diversity of components (i.e. species)
present in a community, and (ii) "vertical diversity" which captures diversity
of components across functional groups, such as the amount trophic categories,
often termed guilds or levels [@allan_stream_2007; @loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017]. Vertical diversity has been
primarily quantified as food chain length [@loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017] but it can also be characterized
by the food web structure [@wang_biodiversity_2017] such as connectance and
average trophic level. Horizontal and vertical diversity are, however, not
independent from each other. Connectance is known to covary negatively with
species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010].

**Horizontal and vertical diversity are, however, not independent from each other
[@duffy_functional_2007; @zhao_horizontal_2019]. The add or removal of one
species (i.e. ) can have different effects depending on its vertical position in the
trophic chain [@duffy_functional_2007]. Hence, it is crucial to study their
interplay to improve our understanding of ecosystem dynamics.**

@@

Horizontal diversity has been shown to promote higher biomass production in a
variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase primary productivity [@cardinale_species_2002,
@tilman_productivity_1996; @grace_integrative_2016]. Two mechanisms have been
proposed to rise this effect [@loreau_partitioning_2001]. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
Positive select effect arise when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
As for total biomass, empirical studies have generally demonstrated positive
relationships with species richness in plant communities [@hector_plant_1999;
@tilman_biodiversity_2014; but see @grace_integrative_2016] and mesocosms
[@pennekamp_biodiversity_2018]. When adding the vertical diversity, e.g.
consumers species, @gamfelt_species_2005 and @thebault_relationship_2006 have found that total
biomass increased with increasing species diversity at two trophic levels.
But this positive relation seems to strongly depend on the niche differentiation of
the consumer species, i.e. their resource complementarity,
@thebault_relationship_2006. Hence, the total biomass-species richness
relationship seems to be supported by the same mechanisms within consumers and
within plants. (also duffy_ecosystem_2005)

The effect of the vertical diversity itself does not seem to affect total
biomass. But seek food chain length, connectance

You should connect average trophic level to biomass repartition (good when lots
of omnivory and no trophic chain)

- Effect of food web structure on total biomass are relatively absent

However, increasing consumer species richness
results in higher consumer biomass to the detriment of primary producer biomass
[@gamfelt_species_2005; @thebault_relationship_2006].

- Horizontal div at different trophic levels (v & h )
- Vertical div:

The food web structure, vertical diversity, also modulates total biomass and
productivity [@thebault_food-web_2003; @thebault_trophic_2005;
@ogorman_multiple_2012; @dossena_warming_2012; @schneider_animal_2016;
@gamfelt_species_2005]. Both in theoretical and experimental study,

The addition of trophic levels, consumer usually
breaks the positive relationship between plant biomass and plant diversity as a
significant amount of biomass is transfered to consumers
[@thebault_food-web_2003].

It has been shown both in experiments
[@ogorman_multiple_2012; @dossena_warming_2012]
and theoretical models [@thebault_food-web_2003; @schneider_animal_2016;
@wang_biodiversity_2017].

Theoretical models @thebault_relationship_2006
have shown with a theoretical model that total biomass should increase with the
increase species diversity at two trophic level *+[]*. However, the joint effects of horizontal and vertical
diversity on total biomass has been rarely tested, especially in natural
communities. It is not clear whether horizontal and vertical diversity just
change the distribution of biomass across trophic level or it increases the
total amount of  biomass (*It misses a lot of references here*).

@@

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increase temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018]. While in
natural communities, same relation has been found for temporal biomass stability
in stream fish communities [@franssen_annual_2011] and temporal abundance
stability of birds and butterflies [@olivier_independent_nodate, but not bats].
Two mechanisms are involved into temporal stability of communities, (1)
synchrony and (2) species variability [@loreau_species_2008;
@thibaut_understanding_2013]. Synchrony measures the temporal covariation of
species. Very low synchrony indicates compensatory temporal variation between
species and so increased temporal stability. Species richness has been generally
found to decrease synchrony between species [@loreau_species_2008;
@thibaut_understanding_2013; @bluthgen_land_2016; @olivier_independent_nodate,
but @declerck_species_2006]. The average temporal variability of species is
measured as a average coefficient of variation of species, weighted by species
abundance/biomass. Higher species temporal variabiliy thus results in lower
temporal stability. Positive [@tilman_biodiversity_2006], neutral
[@jiang_different_2009; @olivier_independent_nodate] and negative
[@franssen_annual_2011; @olivier_independent_nodate] relationships between
species variability and species richness have been found. This different
patterns can be dependent on species mean variance relationships and on the
relationship between evenness and species richness
[@thibaut_understanding_2013].  Synchrony has been found to be the main driver
of temporal stability [@tilman_biodiversity_2006; @olivier_independent_nodate]
and responsible for the positive relationship between horizontal diversity and
temporal stability of biomass and abundance. (*Should I add more details about
the mecanisms?*)

Vertical diversity, represented by network structure, has been central to the
literature of diversity-stability debate and largely explored by theoretical
ecology [@may_will_1972; @angelis_stability_1975;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. Because species
interactions make species dynamics interdependant, they have a tremendous role
on stability notably through its effects on the propagation of perturbations.
However, it is important to note that theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the number
of secondary extinctions after species removal [@duffy_why_2009]. Those
stability metrics hardly relates to empirical temporal stability [@arnoldi_resilience_2016].
However, @thebault_trophic_2005 shows for a bi-trophic community
that asynchrony consistently increase temporal stability and a positive
relationship between temporal stability and species richness. However asynchrony
can also increase species variability through interspecific competition, which
amplifies species biomass fluctuations [@thebault_trophic_2005]. Furthermore
species interactions, considered as competition strength and niche breath of
consumers can results in negative or positive temporal stability-species
richness relationships
[@thebault_trophic_2005] depending on the food web configuration. Vertical
diversity, as network structure, can have strong effects on temporal stability.
However, there are few assesments of the effects of horizontal and diversity on
temporal stability in natural communities.

**Justify that horizontal and vertical diversity are linked**
Because horizontal & vertical diversity are not indedepent, they should be
analyzed jointly. For example, connectance generally covaries negatively with
species richness [@dunne_network_2006; @thebault_stability_2010; but see
@winemiller_must_1989]. This is also true for other metrics [@dunne_network_2006;
@delmas_analyzing_2018]. **TO DEVELOP, MOVE or RM**

#### Environment plays a big role in structuring communities

Environment also largely structure communities and is important to take in
account when considering natural communities. Environment directly affects
organism performance, and subsequently, abundance of species and in turn
determines total biomass [Add general ref; @grace_integrative_2016] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @friedpetersen_drivers_2020]. Hydrology of streams
constraints the composition of fish communities [REF toulouse à balle], and has
already been related to fish population temporal stability [@franssen_annual_2011].
However it is not clear which is the relative strength of direct and indirect
effects of environment on total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

@@

The joint effects of horizontal & vertical diversity on total biomass and its
temporal stability remains yet poorly documented, especially in natural
communities. In this study, we questionned (1) whether the relationships between
horizontal diversity and total biomass/temporal stability are positive or
negative in aquatic ecosystems (Figure \@ref(fig:matmet) A). Then, as indicated
by theoretical studies, (2) what are the effects of vertical diversity on total
biomass and its temporal stability? Considering both direct & indirect effects,
(3) What is the magnitude of environmental effects on total biomass and its
temporal stability compare to horizontal and vertical diversity? (4) Which are
the pathways of the effects of environment, horizontal & vertical diversity on total
biomass and its temporal stability ? (4a) For environment, is it mainly (i)
direct, indirect (ii) through horizontal diversity, (iii) through horizontal &
vertical diversity or (iv) indirect through only vertical diversity? (4b) For
horizontal diversity, is it mainly (i) direct, or (ii) indirect through vertical
diversity?  Finally for temporal stability, (4c) how do environment, horizontal
& vertical diversity affect synchrony & species stability ?

We used an standardized monitoring of fish communities in french streams. 99
sites were followed more than 10 years on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

- Q1:  horizontal diversity-biomass horizontal diversity-stability in aquatic systems
- Q2:  in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3: Relative contribution of environment and community structure to biomass
  and stability

@@

@@

Fish communities were monitored across streams of France on the period 1995-2018
by a French governmental agency (Office Français de la Biodiversité), using
electric fishing. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling is done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temperal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more (i.e 10 sampling events or more). The temporal stability of
biomass, computed as inverse of the CV, can cover very different dynamics such
as decreasing, stable or increasing. However, the concept of community
variability referred to variability around an equilibrium, which is rarely
observed or difficult to characterize in natural communities. We finally kept
only the 99 stations (Figure \@ref(fig:matmet) B) which had no trends in
biomass (method detailed in Supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results).

@@

The average trophic level was positively linked to total biomass. This result is
straightforward. As feeding interactions were partly determined by body size and
that body mass increase with body size, it is expected that the more the average
biomass is located at higher trophic level, the more the total biomass increase.
In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance should translate in higher resource transfer
efficiency between prey and predators and so results in higher total biomass. A
reason for such an absence of an effect might be the absence of an interaction
strength knowledge/estimation. A consequence can be that connectance contains
few too informations. [ADD @thebault_relationship_2006].

>>>>

Biodiversity can be quantified in numerous ways: either as the number of species
(i.e.  species richness) and in number of functional groups, which are usually
defined based on the trophic ecology of species, namely the way they acquire
food [@allan_stream_2007]. Those metrics can be organized into (i) "horizontal
diversity", which captures the overall diversity of components (i.e. species)
present in a community, and (ii) "vertical diversity" which captures diversity
of components across functional groups, such as the amount trophic categories,
often termed guilds or levels [@allan_stream_2007; @loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017]. Vertical diversity has been
primarily quantified as food chain length [@loreau_population_2010;
@zhao_horizontal_2019; @wang_biodiversity_2017] but it can also be characterized
by the food web structure [@wang_biodiversity_2017] such as connectance and
average trophic level. Horizontal and vertical diversity are, however, not
independent from each other. Connectance is known to covary negatively with
species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010]. Species diversity also result in modification of
biomass distribution across the food web [@thebault_relationship_2006;
@duffy_ecosystem_2005; @duffy_functional_2007; @schneider_animal_2016].
Hence, it is crucial to study if horizontal and vertical diversity have synergic
or antagonistic effects and how their interplay modulates ecosystem properties.

@@

Horizontal diversity has been shown to promote higher biomass production in a
variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase primary productivity [@cardinale_species_2002,
@tilman_productivity_1996; @grace_integrative_2016]. Two mechanisms have been
proposed to rise this effect [@loreau_partitioning_2001]. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
Positive select effect arise when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
As for total biomass, empirical studies have generally demonstrated positive
relationships with species richness in plant communities [@hector_plant_1999;
@tilman_biodiversity_2014; but see @grace_integrative_2016] and mesocosms
[@pennekamp_biodiversity_2018]. When adding the vertical diversity, e.g.
consumers species, @gamfelt_species_2005 and @thebault_relationship_2006 have found that total
biomass increased with increasing species diversity at two trophic levels.
But this positive relation seems to strongly depend on the niche differentiation of
the consumer species, i.e. their resource complementarity,
@thebault_relationship_2006. Hence, the total biomass-species richness
relationship seems to be supported by the same mechanisms within consumers and
within plants. @schneider_animal_2016 also showed in a allometric food web model
that species richness should increase total animal biomass.
Vertical diversity also affect total biomass. Theoretical studies showed that food
web properties, such as interaction distribution and strength, affect the
vertical distribution of biomass [@thebault_relationship_2006;
@schneider_animal_2016; @barbier_pyramids_2019], i.e. average trophic level.
More precisely, an increase of animal diversity results in
higher animal biomass as more biomass is transfered from plant to animals
[@schneider_animal_2016; @thebault_food-web_2003]. However, the role of vertical
diversity on the total biomass remains unclear in
natural communities.

@@

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increases the temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], and mesocosms [@pennekamp_biodiversity_2018]. The
same relation was found in natural settings for temporal biomass stability in
stream fish communities [@franssen_annual_2011] and temporal abundance stability
of birds and butterflies [@olivier_independent_nodate, but not bats].  Two
mechanisms may govern the temporal stability of communities, (1) species
synchrony and (2) variability [@loreau_species_2008;
@thibaut_understanding_2013]. Species synchrony measures the temporal
covariation of species where low synchrony indicates that species temporal
variation compensate each other at the community level, thereby increasing
temporal stability. Species richness was found to decrease
synchrony between species [@loreau_species_2008; @thibaut_understanding_2013;
@bluthgen_land_2016; @olivier_independent_nodate, but @declerck_species_2006].
The species variability is measured as a average coefficient
of variation of abundance/biomass, weighted by species abundance/biomass. Higher species
temporal variability should thus result in lower temporal stability. Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. This conflicting patterns are dependent on
various mechanisms [see @thibaut_understanding_2013]. Species synchrony has been found to be
the main driver of temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate] and responsible for the positive relationship
between horizontal diversity and temporal stability of biomass and abundance.

Vertical diversity, captured by network structure, has been at the heart of the
diversity-stability debate and largely explored by theoretical ecology
[@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. As species
interactions introduce inter-dependencies in species dynamics, they have a
substantial impact on stability. In particular, the strength of predation and
degree of omnivory in food webs have been shown to favor stability
[@neutel__2007]. Furthermore, competition strength and niche breath of
consumers, can lead to negative or positive temporal stability-species richness
relationships [@thebault_trophic_2005] depending on the food web configuration
[@neutel__2007].  However, it is important to note that theoretical ecology
generally defines stability based on community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the amount
of secondary extinctions following species removal [@duffy_why_2009]. In
practice, these stability metrics hardly relate to empirical temporal stability
[@arnoldi_resilience_2016].

#### Problematic

Overall, the evidence suggests that both a higher horizontal and vertical
diversity should lead to a greater total biomass and stability in communities.
The effects of horizontal and vertical diversity are potentially conflicting and
interacting, as they rely on inter-dependent mechanisms, in such a way that
they need to be addressed jointly. Yet, there is no direct empirical evidence to
date for an interaction between effects of horizontal and vertical diversity on
total biomass and stability in natural communities.

#### Environment can mitigate community structure and ecosystem properties

The environment affects organism performance directly, and thereby, species
abundances and total biomass [Add general ref; @grace_integrative_2016] and
temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @friedpetersen_drivers_2020].
For instance, the hydrology of streams determines the composition of fish
communities, and is also related to fish population temporal stability
[@franssen_annual_2011].  Temperature has potentially conflicting effects on
horizontal and vertical diversity, as higher temperature favour primary
production and fluxes of biomass [@dossena__2012; @ogorman__2012], thus
potentially supporting more species and trophic levels [@arim__2007;
@stegen__2012], but selects against larger body sized individuals
[@dossena__2012; @edeline__2013; @hattab__2016], and thereby should reduce
trophic diversity [@hattab__2016]. It is thus unclear whether these temperature
effects translate indirectly into changes in total standing biomass and
stability in natural food webs. The relative strength of direct and indirect
effects of environment on total biomass and its temporal stability remains an
open question.

- TODO Add on ecosystem size (cf elisa)

@@

Overall, the joint effects of horizontal and vertical diversity on total biomass
and its temporal stability remains poorly documented, especially in natural
communities. In this study, we identified (1) the effects of horizontal
diversity and (2) vertical diversity on total biomass and temporal stability
(Figure \@ref(fig:matmet) A).  Considering both direct & indirect effects, (3)
we quantify if the magnitude of environmental effects on total biomass and its
temporal stability and compare them to that of horizontal and vertical
diversities.  (4) We identify the pathway for the effects of environment,
horizontal and vertical diversities on total biomass and its temporal stability.
(4a) Whether environmental effects operate mainly (i) directly, indirectly (ii)
through horizontal diversity, (iii) through horizontal and vertical diversities
or (iv) indirectly through vertical diversity only? (4b) Whether effects of
horizontal diversity arise (i) directly, or (ii) indirectly through vertical
diversity? Finally, (4c) whether environmental variables, horizontal and
vertical diversities affect species synchrony and species variability ?

We used the ONEMA dataset consisting in standardized sampling of fish
communities in french streams [@edeline__2013]. 99 sites were sampled for more
than 10 years spanning on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

We show that TODO

@@

## Data collection

@@

Fish communities were monitored across stream sections in France on the period 1995-2018
by the French Office French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing, resulting in over XX individual fish sampled [see
@edeline__2013]. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling was done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temporal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more. The temporal stability of biomass, computed as inverse of the
CV, can cover very different dynamics such as decreasing, stable or increasing.
However, the concept of community variability referred to variability around an
equilibrium, which is rarely observed or difficult to characterize in natural
communities. We finally kept only the 99 stations (Figure \@ref(fig:matmet) B)
which had no trends in biomass (method detailed in Supplementary methods).
However, the results were robust to the relaxation of this hypothesis
(Supplementary results).

@@

The average trophic level was positively linked to total biomass. This result is
straightforward. As feeding interactions were partly determined by body size and
that body mass increase with body size, it is expected that the more the average
biomass is located at higher trophic level, the more the total biomass increase.
*Look at Digel et al. 2011, Riede et al. 2011*

In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance should translate in higher resource transfer
efficiency between prey and predators and so results in higher total biomass. A
reason for such an absence of an effect might be the absence of an interaction
strength knowledge/estimation. A consequence can be that connectance contains
few too informations. [ADD @thebault_relationship_2006].
